---
title: "HW 01"
author: "Megan Hokama"
format: html
toc: true
---
#install packages and download data
library()
install.packages("pacman")
pacman::p_load("tidyverse")
install.packages("devtools")
library("devtools")
devtools::install_github("rstudio-education/dsbox")
install.packages('dsbox')
library(dsbox)
library(tidyverse)

#set theme for ggplot
ggplot2::theme_set(ggplot2::theme_minimal(base_size=11))

#set width of code output
options(width = 65)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7,        #7" width
  fig.asp = 0.618,      #the golden ratio
  fig.retina = 3,       #dpi multiplier for displaying HTML output on retina
  fig.align = "center", #center align figures
  dpi = 300             #higher dpi, sharper image
)



glimpse(accidents)
## 1 - Road traffic accidents in Edinburgh
```{r create ggplot by creating a separate column of day types and creating a plot comparing based on type of day}

day_type_vector <- c(Monday = "Weekday", Tuesday = "Weekday", Wednesday = "Weekday", Thursday = "Weekday", Friday = "Weekday",Saturday = "Weekend", Sunday = "Weekend")

# Use function command
type <- function(day_of_week) 
{return(ifelse(day_of_week %in% names(day_type_vector), day_type_vector[day_of_week], "NA"))
}

# Create new column
accidents$typeday <- sapply(accidents$day_of_week, type)

#Create ggplot with them all in one plot
ggplot(data = accidents, aes(x = time, fill = severity)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ typeday, ncol = 1) + 
  scale_fill_discrete(name="Severity")+
  labs(
    x        = "Time of the day",
    y        = "Density",
    title    = "Number of accidents throughout the day",
    subtitle = "By day of week and severity",
  )
```

```

#Interpretation: There are no fatal accidents on the weekend compared to the weekdays. On both weekends and weekdays, accidents increase in density during the evening hours (16:00-20:00). 

## 2 - NYC marathon winners

```{r Comparing Histograms and Boxplots to determine marathon time and gender}
install.packages("openintro")
library("openintro")

#Part A
nyc <- nyc_marathon
library(psych)
describe(nyc)

#Histogram with distribution of marathon times of all runners 
nyc%>%
  ggplot(aes(x=time_hrs)) + 
  geom_histogram() +
  labs (x = "Marathon Time",
        y = "Runner Count")

#Boxplot with distribution of marathon times of all runners
nyc%>%
  ggplot(aes(x=time_hrs)) +
  geom_boxplot() +
  labs(x = "Marathon Time",
      y = "Runner Count")
  
#The histogram gives more information in regards to the count for all time hours. 

#Part B
ggplot(nyc, aes(x=division, y=time_hrs, fill=division)) + 
geom_boxplot() + 
scale_fill_manual(values = c("Men" = rgb(0.1,0.2,0.3), "Women" = rgb(0.1,0.2,0.3))) +
theme(legend.position="none")
labs(x = "Gender",
      y = "Marathon Times") 
      
    #Based on the box plots, men have lower marathon times compared to women. Additionally, it shows women having more variation and outliers in terms of marathon times compared to men. 

#Part C
#The redundancy is the gender legend as it is not needed since the x-axis already indicates this. 
ggplot(nyc, aes(x=division, y=time_hrs)) + 
geom_boxplot() + 
theme(legend.position="none")
labs(x = "Gender",
      y = "Marathon Times") 

#Part D -It's not running and keeps giving an error in recognizing geom_line
ggplot(nyc, aes(x=year, y=time_hrs, shape=division))+
  geom_line() +
  scale_color_manual(values = c("Men" = rgb(0.1,0.2,0.3), "Women" = rgb(0.1,0.2,0.3)))+
  scale_shape_manual(values = c("Men" = 5, "Women" = 6)) +
  scale_fill_discrete(name="Gender") +
  labs(x = "Year", y = "Marathon Time)
```

## 3 - US counties


## 4 - Rental apartments in SF


## 5 - Napoleon’s march.
