---
title: "HW 01"
author: "Megan Hokama"
format: html
toc: true
---
#install packages and download data
install.packages("pacman")
pacman::p_load("tidyverse")
install.packages("devtools")
library("devtools")
devtools::install_github("rstudio-education/dsbox")
install.packages('dsbox')
library(dsbox)

#set theme for ggplot
ggplot2::theme_set(ggplot2::theme_minimal(base_size=11))

#set width of code output
options(width = 65)

# set figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 7,        #7" width
  fig.asp = 0.618,      #the golden ratio
  fig.retina = 3,       #dpi multiplier for displaying HTML output on retina
  fig.align = "center", #center align figures
  dpi = 300             #higher dpi, sharper image
)



glimpse(accidents)
## 1 - Road traffic accidents in Edinburgh
```{r identify day type using if-else command}

day_type_vector <- c(Monday = "Weekday", Tuesday = "Weekday", Wednesday = "Weekday", Thursday = "Weekday", Friday = "Weekday",Saturday = "Weekend", Sunday = "Weekend")

# Use function command
type <- function(day_of_week) 
{return(ifelse(day_of_week %in% names(day_type_vector), day_type_vector[day_of_week], "NA"))
}

# Create new column
accidents$typeday <- sapply(accidents$day_of_week, type)

#Create ggplot with them all in one plot
ggplot(data = accidents, aes(x = time, fill = severity)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ typeday, ncol = 1) + 
  scale_fill_discrete(name="Severity")+
  labs(
    x        = "Time of the day",
    y        = "Density",
    title    = "Number of accidents throughout the day",
    subtitle = "By day of week and severity",
  )
```

```


## 2 - NYC marathon winners

```{r label-me-2}

```

## 3 - US counties


## 4 - Rental apartments in SF


## 5 - Napoleon’s march.
